Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Latest

Imagine five philosopers sitting around a table, each with a plate of food in
front of them and a fork between each two plates.  Each philosopher requires
two forks to eat.  To start eating, a philosopher first picks up the fork on
the left, then the fork on the right.  Each philospher likes to take breaks
from eating to think for a while.  To do so, the philosopher puts down both
forks.  Each philosopher repeat this procedure.  Dijkstra had them repeating
this for ever, but for the purposes of this book, philosphers can leave
the table when they're not eating.

@@

\noindent
There are four conditions that must hold for deadlock to occur~\cite{CES71}:
\begin{enumerate}
\item \emph{Mutual Exclusion}: each resource can only be used by one process at a time:
\item \emph{Hold and Wait}: each process holds resources it already allocated while it
waits for other resources that it needs;
\item \emph{No Preemption}: Resources cannot be forcibly taken back from processes that
allocated them;
\item \emph{Circular Wait}: There exists a directed circular chain of processes, each waiting
to allocate a resource held by the next.
\end{enumerate}

@@

To implement a \emph{No Hold and Wait} solution, a philosopher would need a
way to lock both the left and right forks at the same time.  Locks do not
have such an ability, and neither do semaphores. so we re-implement the
Dining Philosophers using condition variables that allow one to wait for
arbitrary application-specific conditions.

\autoref{fig:dinerscv} demonstrates how this might be done.
We use a single mutex for the diners, and, for each fork, a boolean
and a condition variable.  The boolean indicates if the fork has been
taken.
Each diner waits if either the left or right fork is already taken.
But which condition variable to wait on?
The code demonstrates an important technique to use when waiting for
multiple conditions.
\index{multiple conditions, waiting on}%
The condition in the \textbf{while} statement is the negation of the
condition that the diner is waiting and consists of two disjuncts.
Within the \textbf{while} statement,
there is an \textbf{if} statement for each disjunct.
The code waits for either or both forks if necessary.  After that, it goes
back to the top of the \textbf{while} loop.

>>>>

Imagine five philosopers sitting around a table, each with a plate of food in
front of them and a fork between each two plates.  Each philosopher requires
two forks to eat.  To start eating, a philosopher first picks up the fork on
the left, then the fork on the right.  Each philosopher likes to take breaks
from eating to think for a while.  To do so, the philosopher puts down both
forks.  Each philosopher repeat this procedure.  Dijkstra had them repeating
this for ever, but for the purposes of this book, philosophers can leave
the table when they're not eating.

@@

\noindent
There are four conditions that must hold for deadlock to occur~\cite{CES71}:
\begin{enumerate}
\item \emph{Mutual Exclusion}: each resource can only be used by one process at a time:
\item \emph{Hold and Wait}: each process holds resources it already allocated while it
waits for other resources that it needs;
\item \emph{No Preemption}: Resources cannot be forcibly taken away from processes that
allocated them;
\item \emph{Circular Wait}: There exists a directed circular chain of processes, each waiting
to allocate a resource held by the next.
\end{enumerate}

@@

To implement a \emph{No Hold and Wait} solution, a philosopher would need a
way to lock both the left and right forks at the same time.  Locks do not
have such an ability, and neither do semaphores. so we re-implement the
Dining Philosophers using condition variables that allow one to wait for
arbitrary application-specific conditions.
%
\autoref{fig:dinerscv} demonstrates how this might be done.
We use a single mutex for the diners, and, for each fork, a boolean
and a condition variable.  The boolean indicates if the fork has been
taken.
Each diner waits if either the left or right fork is already taken.
But which condition variable to wait on?
The code demonstrates an important technique to use when waiting for
multiple conditions.
\index{multiple conditions, waiting on}%
The condition in the \textbf{while} statement is the negation of the
condition that the diner is waiting and consists of two disjuncts.
Within the \textbf{while} statement,
there is an \textbf{if} statement for each disjunct.
The code waits for either or both forks if necessary.  After that, it goes
back to the top of the \textbf{while} loop.
